---
title: "R Notebook"
output: html_notebook
---

```{r}
library(rmgarch)
library(parallel)
library(quantmod)
library(e1071) 
library(matlib)
library(tidyr)
library(dplyr)
library(e1071) 
library(tidyverse)
library(tseries)
```


```{r}
#  Describe analysis interval
startDate=as.Date('2016-04-20')
endDate=as.Date('2021-05-15')

#  Stocks
getSymbols( "AAPL", from=startDate,to=endDate )
getSymbols( "GOOG" ,from =startDate, to=endDate)
getSymbols( "TSLA" ,  from=startDate, to=endDate)

rAAPL<-weeklyReturn(AAPL)
rGOOG<-weeklyReturn(GOOG)
rTSLA<-weeklyReturn(TSLA)

#  Fix table
rX<-data.frame(rAAPL ,rGOOG ,rTSLA )
names(rX)[1]<-"rAAPL"
names(rX)[2]<-"rGOOG"
names(rX)[3]<-"rTSLA"

#  Perform kpss test on all stocks
kpss.test(rX[,1], null="Trend")
kpss.test(rX[,2], null="Trend")
kpss.test(rX[,3], null="Trend")
 
par(mfrow=c(3,1)) 
plot(rX[,1])
plot(rX[,2])
plot(rX[,3])
```

Descriptive statistics
```{r}
stats = rX %>% 
  summarize_all(funs(mean, sd, skewness, kurtosis, min, max)) %>%
  gather() %>%
  separate(key, c("var","type")) %>%
  spread(type, value)

dataStats = as.data.frame(t(stats[c(1,5,3,4,6,7,2)]))
colnames(dataStats) = c("APPL","GOOG", "TSLA")

data[1,] = c("APPL","GOOG", "TSLA")
row.names(dataStats)[1] = c("stock")
```


ARMA models for univariate case
```{r}
aic_criterion = matrix(data = NA, nrow = 4, ncol = length(rX))

#aic table to find the best model
for (i in 1:length(rX)) {
  aic_criterion[1,i] = arima(rX[,i], order = c(1,0,1))$aic
  aic_criterion[2,i] = arima(rX[,i], order = c(1,0,2))$aic
  aic_criterion[3,i] = arima(rX[,i], order = c(2,0,1))$aic
  aic_criterion[4,i] = arima(rX[,i], order = c(2,0,2))$aic
}

data.frame(aic_criterion)
```

Mean and standard deviation of portfolio returns.
```{r}
means = matrix(data = NA, nrow = length(rX), ncol = 1)
means[1,1] = arima(rX[,1], order = c(2,0,2))$coef["intercept"]
means[2,1] = arima(rX[,2], order = c(1,0,1))$coef["intercept"]
means[3,1] = arima(rX[,3], order = c(1,0,1))$coef["intercept"]

means
```

Test which univariate GARCH model is the best for every stocks seperately.
```{r}
garch_models = matrix(data = NA, nrow = 4, ncol = length(rX))
for (i in 1:length(rX)) {
  ug_spec = ugarchspec(mean.model=list(armaOrder=c(1,1)))
  ugfit = ugarchfit(spec = ug_spec, data = rX[,i], AIC = TRUE)
  garch_models[1,i] = ifelse(ugfit@fit$convergence == 1, NA , infocriteria(ugfit)[1]) 
  
  ug_spec = ugarchspec(mean.model=list(armaOrder=c(1,2)))
  ugfit = ugarchfit(spec = ug_spec, data = rX[,i], AIC = TRUE)
  garch_models[2,i] = ifelse(ugfit@fit$convergence == 1, NA , infocriteria(ugfit)[1]) 
  
  ug_spec = ugarchspec(mean.model=list(armaOrder=c(2,1)))
  ugfit = ugarchfit(spec = ug_spec, data = rX[,i], AIC = TRUE)
  garch_models[3,i] = ifelse(ugfit@fit$convergence == 1, NA , infocriteria(ugfit)[1]) 
}

data.frame(garch_models)

res1 = residuals(ugfit, standardize = TRUE)
res2 = residuals(ugfit2, standardize = TRUE)
cov(res1, res2, res1)
```

Taking only the best garch fits to get cov matrix
```{r}
#  create different specifications
ug_spec = ugarchspec(mean.model=list(armaOrder=c(2,2)), variance.model = list( garchOrder = c(1, 2)))
ugfit1 = ugarchfit(spec = ug_spec, data = rX[,1], AIC = TRUE)

ug_spec = ugarchspec(mean.model=list(armaOrder=c(1,1)), variance.model = list( garchOrder = c(1, 1)))
ugfit2 = ugarchfit(spec = ug_spec, data = rX[,2], AIC = TRUE)

ug_spec = ugarchspec(mean.model=list(armaOrder=c(1,1)), variance.model = list( garchOrder = c(1, 2)))
ugfit3 = ugarchfit(spec = ug_spec, data = rX[,3], AIC = TRUE)

#  follow standardized residuals
res1 = residuals(ugfit, standardize = TRUE)
res2 = residuals(ugfit2, standardize = TRUE)
res3 = residuals(ugfit3, standardize = TRUE)

sigma1 = as.numeric(sigma(ugfit)[265])
sigma2 = as.numeric(sigma(ugfit2)[265])
sigma3 = as.numeric(sigma(ugfit3)[265])

omega_uni[1,1] = as.numeric(cor(res1, res1)*sigma(ugfit1)[265] *sigma(ugfit1)[265])
omega_uni[2,2] = as.numeric(cor(res2, res2)*sigma(ugfit2)[265] *sigma(ugfit2)[265])
omega_uni[3,3] = as.numeric(cor(res3, res3)*sigma(ugfit3)[265] *sigma(ugfit3)[265])

omega_uni[2,1] = omega_uni[1,2] = as.numeric(cor(res2, res1)*sigma(ugfit2)[265] *sigma(ugfit1)[265])
omega_uni[3,1] = omega_uni[1,3] = as.numeric(cor(res3, res1)*sigma(ugfit3)[265] *sigma(ugfit1)[265])
omega_uni[3,2] = omega_uni[2,3] = as.numeric(cor(res3, res2)*sigma(ugfit3)[265] *sigma(ugfit2)[265])

dt = data.frame(res1, res2, res3)

#  metrics of every suggested model specification
cov(dt)
mean(means)
```

Find correlation matrix and its inverse to calculate out 
```{r}
#  mean-variance maximization algorithm (using separate specified models for each stock)
omega_uni = cov(dt)
omega.inv_uni =  inv(omega_uni)

e = matrix(rep(1, length(rX)), nrow = length(rX), ncol = 1)

#  tau can be set manually if you already noticed a pattern
tau = seq(2.86,2.94, by = 0.01)

#  create a matrix for weights
weights_uni = matrix(data = NA, nrow = length(tau), ncol =length(rX) )

#  to follow assumptions of normality of returns
Z = qnorm(0.05, mean = 0, sd = 1)

#  iterative maximization algorithm
for (i in 1:length(tau)) {
A = t(e) %*% omega.inv_uni %*% e
  
B = (2*tau[i] + 1) %*% (t(means) %*% omega.inv_uni %*% e + t(e) %*% omega.inv_uni %*% means)

C = (2*tau[i] + 1)^2 %*% (t(means) %*% omega.inv_uni %*% means) + Z
 
lambda = (-B + sqrt(B^2 - 4%*%A%*%C)) %*% 1/A

weights_uni[i,]  = ((2*tau[i] + 1) * omega.inv_uni %*% means + lambda[[1]] * omega.inv_uni %*% e) /
        ( (2*tau[i] + 1) * t(e) %*% omega.inv_uni %*% means + lambda[[1]] * t(e) %*% omega.inv_uni %*% e)[[1]]
}

data.weights = data.frame(tau, weights_uni)
data.weights_uni = data.weights
```

Apply DCC-GARCH model to rX dataframe and get dynamical covariation matrix.
```{r}
#  fit all models into "one"
uspec.n = multispec(replicate(length(rX), ugarchspec(mean.model = list(armaOrder = c(1,1)), 
                                                     variance.model = list(garchOrder = c(1,1), 
                                                model = "sGARCH"))))

multf = multifit(uspec.n, rX)
spec1 = dccspec(uspec = uspec.n, dccOrder = c(1, 1), distribution = 'mvnorm')
fit1 = dccfit(spec1, data = rX, fit.control = list(eval.se = TRUE), fit = multf, AIC = TRUE)

cov1 = rcov(fit1)  # extracts the covariance matrix
```


```{r}
omega_dcc = cov1[,,dim(cov1)[3]]
omega.inv_dcc = inv(omega_dcc)

tau = seq(0, 120, by=0.1)

weights_dcc = matrix(data = NA, nrow = length(tau), ncol =length(rX) )

for (i in 1:length(tau)) {
A = t(e) %*% omega_dcc %*% e
  
B = (2*tau[i] + 1) %*% (t(means) %*% omega.inv_dcc %*% e + t(e) %*% omega.inv_dcc %*% means)

C = (2*tau[i] + 1)^2 %*% (t(means) %*% omega.inv_dcc %*% means) + Z
 
lambda = (-B + sqrt(B^2 - 4%*%A%*%C)) %*% 1/A

weights_dcc[i,]  = ((2*tau[i] + 1) * omega.inv_dcc %*% means + lambda[[1]] * omega.inv_dcc %*% e) / ( (2*tau[i] + 1) * t(e) %*% omega.inv_dcc %*% means + lambda[[1]] * t(e) %*% omega.inv_dcc %*% e)[[1]]
}

data.weights_dcc = data.frame(tau, weights_dcc)

# 0<=tau <=4.70 hence we can remove redundant values to plot it more clearly
data.weights_dcc = data.weights_dcc[c(8:32),]
weights_dcc = weights_dcc[c(8:32),]
```

Mean - VaR optimization using typical model and DCC
```{r}
VaR = matrix(data = NA, nrow = dim(data.weights_uni)[1], ncol = 1)
mean_uni.after = weights_uni %*% means

for (i in 1:dim(data.weights_uni)[1]) {
  VaR[i,1] = -3 *(weights_uni[i,] %*% means + Z *sqrt(weights_uni[i,] %*% omega_dcc %*% as.matrix(t(weights_uni)[,i])))
}

#  DCC case
VaR_dcc = matrix(data = NA, nrow = dim(data.weights_dcc)[1], ncol = 1)
mean_dcc.after =  weights_dcc %*% means

for (i in 1:dim(data.weights_dcc)[1]) {
  VaR_dcc[i,1] = -3 *(weights_dcc[i,] %*% means + Z *sqrt(weights_dcc[i,] %*% omega_dcc %*% as.matrix(t(weights_dcc)[,i])))
}
```

Random plots to analyze the optimal var-mean, with what coefficients (or distribution of portfolio) it becomes optimal
```{r}
tau = seq(2.86,2.94, by = 0.01)
par(mfrow=c(1,2)) 

plot(tau, mean_uni.after %*% 1/VaR,  type = "b", pch = 25, bty="l", ylab = "Mean VaR ratio")
abline(v = 2.88, col="red", lwd=3, lty=2)

plot(VaR, mean_uni.after, type = "b", pch = 25, bty="l", ylab = "Mean")
tau = seq(0.7,3.1, by = 0.1)

plot(tau, mean_dcc.after %*% 1/VaR_dcc, type = "b", pch = 25, bty="l", ylab = "Mean VaR ratio")
abline(v = 2.3, col="red", lwd=3, lty=2)
plot(VaR_dcc,mean_dcc.after, type = "b", pch = 25, bty="l", ylab = "Mean")

#  DCC case
jpeg("plotM.jpeg", width = 480, height = 480) # height and width can choose as your wish 
par(mfrow=c(1,2)) 

plot(tau, mean_dcc.after %*% 1/VaR_dcc, type = "b", pch = 25, bty="l", ylab = "Mean VaR ratio")
abline(v = 2.3, col="red", lwd=3, lty=2)

plot(VaR_dcc,mean_dcc.after, type = "b", pch = 25, bty="l", ylab = "Mean")
dev.off()
```


